# Exercise 4.14

## Problem Statement

Using the `Students` data file, construct a 95% confidence interval:

-   **a**: For the mean weekly number of hours spent watching TV.
-   **b**: To compare females and males on the mean weekly number of hours spent watching TV.

## Solution

### Load Data

```{r}
library(dplyr)
students_data <- read.table("Students.dat.txt", header = TRUE)
head(students_data)
```

### a - 95% Confidence Interval for Mean Weekly Hours Watching TV

```{r}
mean_tv <- mean(students_data$tv, na.rm = TRUE)
sd_tv <- sd(students_data$tv, na.rm = TRUE)
n_tv <- sum(!is.na(students_data$tv))

error_margin <- qt(0.975, df = n_tv - 1) * (sd_tv / sqrt(n_tv))
ci_lower <- mean_tv - error_margin
ci_upper <- mean_tv + error_margin

cat("95% Confidence Interval for mean weekly hours watching TV: [", ci_lower, ",", ci_upper, "]\n")
```

### b - Comparison of Mean Weekly Hours Watching TV Between Females and Males

```{r}
females <- students_data %>% filter(gender == 1) %>% select(tv) %>% unlist()
males <- students_data %>% filter(gender == 0) %>% select(tv) %>% unlist()

mean_female <- mean(females, na.rm = TRUE)
sd_female <- sd(females, na.rm = TRUE)
n_female <- sum(!is.na(females))

mean_male <- mean(males, na.rm = TRUE)
sd_male <- sd(males, na.rm = TRUE)
n_male <- sum(!is.na(males))

se_diff <- sqrt((sd_female^2 / n_female) + (sd_male^2 / n_male))
mean_diff <- mean_female - mean_male

error_margin_diff <- qt(0.975, df = min(n_female, n_male) - 1) * se_diff
ci_lower_diff <- mean_diff - error_margin_diff
ci_upper_diff <- mean_diff + error_margin_diff

cat("95% Confidence Interval for the difference in mean weekly hours watching TV (Females - Males): [", ci_lower_diff, ",", ci_upper_diff, "]\n")
```

the 95% confidence interval for the difference in mean weekly hours spent watching TV between females and males is [-2.09, 5.06]. Since this interval includes zero, means that there may not be a significant difference between the two groups.
